---
title: "Bookbinders Case Study"
author: "Austin Vanderlyn, Christine Kelly, Richard Tarbell"
date: "2/16/2022"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(logistf)
library(tidyverse)
library(car)
library(broom)
library(DescTools)
library(ROCR)
library(ResourceSelection)
library(e1071)
```


```{r, include=FALSE}
set.seed(100)
BBBC_Train <- read_excel("Data/BBBC-Train.xlsx",col_names = TRUE)
BBBC_Test <- read_excel("Data/BBBC-Test.xlsx",col_names = TRUE)
```


### I. Executive Summary 
In an effort to find the best predictions for which customers would purchase the book, The Art History of Florence, we ran three different models.  The model we felt performed the best was the logistic regression model.  This decision was based on the results seen in the confusion matrix, which showed the highest number of predicted customers who purchased the book.  The linear regression model was unsuccessful since the variable of interest is binary and not continuous.   The Support Vector Machine (SVM) was tested with three different hyper parameters, but was still not able to perform as well as the logistic model due to the non purchasing group being over represented in the data set.  

In the future using the either the logistic model or the SVM with data that is more balanced would provide the best results in finding which customers to target for mail campaigns.  Using these models provides a much greater profit for the Bookbinders Book Club (BBBC).  These models would also be able to be customized for different regions and campaigns. 


### II. The Problem - Austin

### III. Review of Related Literature 

Linear Regression is a parametric model used to find if any explanatory variables influence the response variable.  There are four assumptions for linear regression.  The first is the linear relationship between X and Y.  The second is normal distribution.  The third is homoscedasticity.  Finally, all observations must be independent of each other.  The model finds the best line to predict the behavior of Y, when X is increased by 1 unit.  Y must be continuous for the model to work.

Logistic Regression is used to predict the probability of an outcome based on given variables, or to see how variables are related to the outcome.  For the logistic model Y is binary and not continuous.  Interpretation of the logistic model is not as straightforward and requires an understanding of odds and odds ratio.  

Support Vector Machine (SVM) is a popular model that was developed in the 1990s by Vladimir Vapnik and Corinna Cortes.  This model finds a hyperplane that separates the data as well as possible and allows some misclassifications. To accommodate a non-linear boundary between classes, SVM enlarges the feature space using polynomial terms.  The SVM enlarges this feature space, using kernel tricks, in a way that is efficient with these computations. 

All three of these models are beneficial in providing insights and predictions when applied to marketing campaign selections.

### IV. Methodologies

The first model attempted was a linear regression.  This model is not useful for this dataset because the response variable in linear regression must be continuous.  In this case  the response variable Choice is categorical and must be converted to a factor.  If the regression model is used leaving the Choice as numeric, the information it provides is not useful for the criteria we are trying to meet.  





Along with the Linear and Logit model we performed tests using the SVM models. Some advantages of using SVM are its versatility when working with high dimensional spaces and it's ability to be memory efficient. One issue we ran into with the SVM was it's poor ability to handle class imbalance.The SVM model used for the BookBinders took two hyperparameters as input, gamma and cost. The gamma hyperparameter is used to give curvature weight of the decision boundary or how far the influence of a single training sample reaches. The cost hyperparameter is used to control regularization which tells the algorithm how much you wish to misclassify each training sample. Testing the SVM models we have various kernels we can use Radial, Linear, Polynomial, and Sigmoid. In order to reduce overfitting we start with the simplest model which is the linear SVM and worked our way up to polynomial excluding the Sigmoid for now. Using a Train/Test split on the Training sample we were able to narrow down a model we were that was able to produce high Specificity values on the Test data.





### V. Data / Cleaning - Richard

The datasets required minimal cleaning. Both the Train (`BBBC_Train`) and Test (`BBBC_Test`) datasets originally had 12 variables and the Training set consisted of 1600 observations while the Testing set had 2300. Performing basic analysis of the data we saw there were no missing values and there happened to be a column in each titled `Observation`. We removed this column because it would add no significant value to our models.

Checking for correlations within the pairwise plot the strongest correlation came out to be between `Last_Purchase` and `First_Purchase` which resemble the months since the customers last purchase and the months since their first purchase. With this being the only positive correlation > 0.7 we may run into the issue of multicollinearity. 

The variables `Choice` and `Gender` were both treated as numeric variables within the data however we converted these to factor variables given their binary values. Further exploring these variables we discovered that approximately 70% of customers who did not a purchase were Males and \~54% of the customers who did make a purchase were also Males. Comparing the customers who did not make a purchase to those who did we find a class imbalance with only 25% of the `Choice` class to being customers who did make a purchase. This may cause issues in model accuracy given that SVMs do not perform well on imbalanced datasets.





### VI. Findings 

The linear regression is not able to provide useful information for this problem since the variable of interest is binary and not continuous.  

Using a 3 different kinds of model in determining what parameters lead to customers purchasing a certain book. This was important because the Bookbinders Book Club doesn't want to waste money sending out mailings to customers who will not end up purchasing a book. With our three models we tested Linear Regression, Logistic Regression, and Support Vector Machines.

Before we performed logistic regression we removed `Last_Purchase` and `First_Purchase` from the model due to their high collinearity. After this we performed a step wise selection which allowed us to remove the `P_Youth` predictor as it was insignificant. This left us with a final model of 
$\log\frac{1}{(1-p)})$ = -0.289 + 1.244\*`P_Art` - 0.088\*`Frequency` - 0.812\*`Gender1` - 0.294\*`P_Cook` - 0.282\*`P_DIY` + 0.002\*`Amount_Purchased` - 0.196\*`P_Child`. 

Using the equation for logistic regression and taking the exponential for each coefficient value we can state the following,

The odds of a customer buying The Art History of Florence change by a factor of 3.46 with each additional art book purchased, assuming other variables remain constant.

The odds of a customer buying The Art History of Florence change by a factor of 0.915 with each additional book purchased, assuming other variables remain constant.

The odds of a male customer are .443 times that of a female customer, assuming other variables remain constant.

The odds of a customer buying The Art History of Florence change by a factor of 0.745 with each additional cook book purchased, assuming other variables remain constant.

The odds of a customer buying The Art History of Florence change by a factor of 1.002 with each additional dollar spent, assuming other variables remain constant.

The odds of a customer buying The Art History of Florence change by a factor of 1.21 with each additional children’s book purchased, assuming other variables remain constant.


**Logistic Regression with 0.5 cutoff**

\begin{center}
\begin{tabular}{|c || c c|} 
 \hline
   & Reference &   \\ [0.5ex] 
 \hline\hline
 Prediction & 0 & 1 \\ 
 \hline
 0 & 1986 & 110 \\
 \hline
 1 & 131 & 73 \\
 \hline
\end{tabular}
\end{center}



**Logistic Regression with optimal cutoff (0.23)**

\begin{center}
\begin{tabular}{|c || c c|} 
 \hline
   & Reference &   \\ [0.5ex] 
 \hline\hline
 Prediction & 0 & 1 \\ 
 \hline
 0 & 1490 & 606 \\
 \hline
 1 & 55 & 149 \\
 \hline
\end{tabular}
\end{center}



Following the logistic regression was the test of the Support Vector Machine method using Linear, Radial, and Polynomial Kernels. A wide range of hyperparameters were used for these kernels. For the Linear and Radial kernels we used gamma values with a range defined by the sequence `seq(.01, .1, by = .01)`, with all three we tested cost with a range defined by the sequence `cost = seq(.1, 1, by = .1)`. Finally for the polynomial kernel the degree sequence was defined by `seq(2, 4, by = 1)` and the tested coef0 values came from an array of numbers  `(0.01, 0.1, 0.25, 0.5, 0.75, 1, 2, 3)`. The following confusion matrices resulted.

**SVM with Linear Kernel**

\begin{center}
\begin{tabular}{|c || c c|} 
 \hline
   & Reference &   \\ [0.5ex] 
 \hline\hline
 Prediction & 0 & 1 \\ 
 \hline
 0 & 2088 & 192 \\
 \hline
 1 & 8 & 12 \\
 \hline
\end{tabular}
\end{center}



**SVM with Radial Kernel**

\begin{center}
\begin{tabular}{|c || c c|} 
 \hline
   & Reference &   \\ [0.5ex] 
 \hline\hline
 Prediction & 0 & 1 \\ 
 \hline
 0 & 2047 & 160 \\
 \hline
 1 & 49 & 44 \\
 \hline
\end{tabular}
\end{center}

**SVM with Polynomial Kernel**

\begin{center}
\begin{tabular}{|c || c c|} 
 \hline
   & Reference &   \\ [0.5ex] 
 \hline\hline
 Prediction & 0 & 1 \\ 
 \hline
 0 & 2018 & 140 \\
 \hline
 1 & 78 & 64 \\
 \hline
\end{tabular}
\end{center}


Comparing the Logistic Regression Model and the SVMs we can view the accuracy's and how well the models predicted false positives or false negatives.


\begin{center}
\begin{tabular}{|c | c c c c||} 
 \hline
  & Logit & SVM-Linear & SVM-Radial & SVM-Polynomial \\ [0.5ex] 
 \hline\hline
 Accuracy & 0.7126 & 0.913 & 0.9091 & 0.9052 \\ 
 \hline
 Specificity & 0.9644 & 0.9962 & 0.9766 & 0.9628 \\
 \hline
 Sensitivity & 0.1974 & 0.0588 & 0.2157 & 0.3137 \\
 \hline
 AUC & 0.787 &  &  &  \\
 \hline
 Cost &  & 0.3 & 0.8 & 1 \\
 \hline
 Degree &  &  &  & 2 \\ 
 \hline
 Coef0 &  &  &  & 0.5 \\ 
 \hline
\end{tabular}
\end{center}




### VII. Conclusion and Recommendations

Performing our analysis we tested three different modeling techniques in order to find the best procedure to predict if a customer would purchase The Art History of Florence. Using Linear Regression, Logistic Regression, and Support Vector Machines we analyzed a training dataset to find the best predictors and hyper parameters to use when predicting a customer and their purchase choice. With the Support Vector Machines (SVM) we tested three different kernels Linear, Radial, and Polynomial. Out of the three the Linear kernel performed the best with accuracy of 91.3%. The SVM-Linear was extremely successful in correctly classifying customers who did not make a purchase, however, it did fall short in successfully predicting customers who did make a purchase. This can be seen in the confusion matrix and by the low sensitivity/high specificity outputted by the model. This performance is likely due to the class imbalance found in the training set which is one of SVM's weaknesses.

In order to determine which customers the campaign should target the Bookbinders Book Club should use the logistic regression model. The logit model was the most successful in predicting which customers did purchase a book. While our logistic regression model did not have the highest accuracy (0.7126) or the best sensitivity/specificity it was was the most successful in predicting customers who did purchase a book as seen in the confusion matrix table.

If we wanted to make our models more accurate with higher sensitivity and specificity values we would potentially need a new sample of the data where the class distribution is more evenly balanced. If the company wanted to develop expertise in any of the models we would recommend logistic regression and SVM. Both of these models were accurate however SVM was not able to perform at it's peak due to the class imbalance. For future modeling efforts we recommend having a simple yet straight forward way to pull data from the database that is evenly distributed with all the predictors required. This can be automated through a dashboard or using SQL queries and a modeling script in R or Python to automatically perform accurate analysis for the company.



### Appendixs


##### Data importing and cleaning


Exploration

Any missing values?
```{r}
anyNA(BBBC_Train)
anyNA(BBBC_Test)
```


Check the size of the training and test datasets
```{r}
dim(BBBC_Train)

dim(BBBC_Test)
```

Class imbalance
```{r}
table(BBBC_Train$Choice)
```

View details of the data
```{r}
str(BBBC_Train)
```



Remove `Observation` variable and convert `Choice` to factor

```{r}
BBBC_Train = subset(BBBC_Train, select = -Observation)
BBBC_Test = subset(BBBC_Test, select = -Observation)

BBBC_Train$Choice = as.factor(BBBC_Train$Choice)
BBBC_Test$Choice = as.factor(BBBC_Test$Choice)
```



```{r, include=FALSE}
# code to get histograms in pair plot
panel.hist <- function(x, ...) {
    usr <- par("usr")
    on.exit(par(usr))
    par(usr = c(usr[1:2], 0, 1.5))
    his <- hist(x, plot = FALSE)
    breaks <- his$breaks
    nB <- length(breaks)
    y <- his$counts
    y <- y/max(y)
    rect(breaks[-nB], 0, breaks[-1], y, col = rgb(0, 1, 1, alpha = 0.5), ...)
    # lines(density(x), col = 2, lwd = 2) # Uncomment to add density lines
}

panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits = digits)[1]
    txt <- paste0(prefix, txt)
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor)
}
```

```{r, fig.height=8, fig.width=12}
pairs.test = subset(BBBC_Train, select = -Choice)
pairs.test = subset(pairs.test, select = -Gender)

pairs(pairs.test,
      upper.panel = panel.cor,
      diag.panel = panel.hist)
```



```{r}
ggplot(data = BBBC_Train,
       aes(x = factor(ifelse(Choice == 1, "Purchase", "No Purchase")), 
           fill = factor(ifelse(Gender == 0, "Female", "Male")))) + 
    geom_bar(alpha =1, color = "black", stat = "count") + 
    scale_fill_manual(values = c("lavender", "orange")) +
    geom_text(aes(label = scales::percent(..count.. / tapply(..count.., ..x.., sum)[as.character(..x..)])), stat = "count", position = position_stack(vjust=0.5)) +
    labs(fill = "Gender", y="", x="")
```






### Linear Regression Model


```{r}
lm.book <- lm(as.numeric(Choice) ~ ., data = BBBC_Train)
summary(lm.book)
```










### Logistic Regression Model


Change gender to factor for logistic regression model
```{r}
BBBC_Train_Logit = BBBC_Train
BBBC_Test_Logit = BBBC_Test

BBBC_Train_Logit$Gender = as.factor(BBBC_Train_Logit$Gender)

BBBC_Test_Logit$Gender = as.factor(BBBC_Test_Logit$Gender)
```



Create initial logistic regression model
```{r}
glm.train = glm(Choice ~ ., data = BBBC_Train_Logit, family = "binomial")
summary(glm.train)
```

Remove variables with high multicollinearity
```{r}
vif(glm.train)

```

```{r}
glm.train = glm(Choice ~ .-Last_purchase, data = BBBC_Train_Logit, family = "binomial")

vif(glm.train)
```

```{r}
glm.train = glm(Choice ~ .-Last_purchase-First_purchase, data = BBBC_Train_Logit, family = "binomial")

vif(glm.train)
```

Build stepwise model
```{r}
glm.null = glm(Choice ~ 1, data = BBBC_Train_Logit, family = "binomial")
glm.full = glm(Choice ~ .-Last_purchase-First_purchase, data = BBBC_Train_Logit, family = "binomial")

glm.step1 = step(glm.null, scope = list(upper = glm.full), direction = "both", test = "Chisq", trace = FALSE)

summary(glm.step1)
```

Goodness of fit
```{r}
hoslem.test(glm.step1$y, fitted(glm.step1), g = 10)
```

Plot
```{r}
plot(glm.step1, which = 4)
```





```{r}
BBBC_Train_Logit$PredProb = predict.glm(glm.step1, BBBC_Train_Logit, type = "response")
BBBC_Train_Logit$PredChoice = ifelse(BBBC_Train_Logit$PredProb >= 0.5, 1, 0)
caret::confusionMatrix(as.factor(BBBC_Train_Logit$Choice), as.factor(BBBC_Train_Logit$PredChoice), positive = "1")
```

The accuracy of this model is not the best, but the positive predictive value, which is what we care about, is pretty good.



Run model with test data;
```{r}
BBBC_Test_Logit$PredProb = predict.glm(glm.step1, BBBC_Test_Logit, type = "response")
BBBC_Test_Logit$PredChoice = ifelse(BBBC_Test_Logit$PredProb >= 0.5, 1, 0)
caret::confusionMatrix(as.factor(BBBC_Test_Logit$Choice), as.factor(BBBC_Test_Logit$PredChoice), positive = "1")
```

The model has better accuracy with the test data, and still has a high positive predictive value. It could get better by calibrating the sensitivity and specificity.


Calculate and plot AUC
```{r}
pred_train = predict(glm.step1, BBBC_Train_Logit, type = "response")
response_train = BBBC_Train_Logit$Choice
predict_train = prediction(pred_train, response_train)
auc_train = round(as.numeric(performance(predict_train, measure = "auc")@y.values),3)
perform = performance(predict_train, "tpr","fpr")
plot(perform, colorize = T, main = "ROC Curve")
text(0.5,0.5, paste("AUC:", auc_train))
```



```{r}
plot(unlist(performance(predict_train, "sens")@x.values),
     unlist(performance(predict_train, "sens")@y.values),
     type = "l",
     lwd = 2,
     ylab = "Sensitivity",
     xlab = "Cutoff",
     main = paste("Maximized Cutoff", "AUC: ", auc_train))

par(new = TRUE)

plot(unlist(performance(predict_train, "spec")@x.values),
     unlist(performance(predict_train, "spec")@y.values),
     type = "l",
     lwd = 2,
     col = "red",
     ylab="",
     xlab="")

axis(4, at=seq(0, 1, 0.2))
mtext("Specificity", side = 4, col = 'red')

min.diff.glm = which.min(abs(unlist(performance(predict_train, "sens")@y.values) - 
                               unlist(performance(predict_train, "spec")@y.values)))
min.x.glm = unlist(performance(predict_train, "sens")@x.values)[min.diff.glm]
min.y.glm = unlist(performance(predict_train, "sens")@y.values)[min.diff.glm]
optimal.glm = min.x.glm

abline(h = min.y.glm, lty = 3)
abline(v = min.x.glm, lty = 3)
text(min.x.glm,0,paste("optimal threshold=",round(optimal.glm,2)), pos = 4)
```
So now we know that the optimal cutoff threshold is 0.23, so we can refit the predictions to optimize the sensitivity and specificity.

Refit predictions and confusion matrix;
```{r}
BBBC_Test_Logit$PredProb = predict.glm(glm.step1, BBBC_Test_Logit, type = "response")
BBBC_Test_Logit$PredChoice = ifelse(BBBC_Test_Logit$PredProb >= 0.23, 1, 0)
caret::confusionMatrix(as.factor(BBBC_Test_Logit$Choice), as.factor(BBBC_Test_Logit$PredChoice), positive = "1")
```


log(1/(1-p)) = -0.289 + 1.244*P_Art - 0.088*Frequency - 0.812*Gender1 - 0.294*P_Cook - 0.282*P_DIY + 0.002*Amount_Purchased - 0.196*P_Child

The most influential covariates then, are P_Art(number of art books purchased), Frequency(total number of purchases), Gender, P_Cook(number of cookbooks purchased), P_DIY (number of DIY books purchased), Amount_Purchased (total money spent), and P_Child.


```{r}
exp(1.244)
exp(-0.088)
exp(-0.812)
exp(-0.294)
exp(0.002)
exp(0.196)
```

### Midwest Mailing Campaign

Data for the proposed mailing campaign;
50,000 customers
cost of mailing = $0.65 / addressee
cost of book = $15
Selling price of book = $31.95
overhead = 0.45*bookcost










### SVM Model

```{r}
BBBC_Train_SVM = BBBC_Train
BBBC_Test_SVM = BBBC_Test
```

```{r}
str(BBBC_Train_SVM)
```



#### Use training split on BBBC_Train

```{r}
# Splitting data into training and testing sets 70/30 Split
set.seed(1)
tr_ind = sample(nrow(BBBC_Train_SVM), 0.7*nrow(BBBC_Train_SVM), replace=FALSE)
book.train.split = BBBC_Train_SVM[tr_ind,]
book.test.split = BBBC_Train_SVM[-tr_ind,]
```


#### SVM with Linear Kernel

```{r}
svm_form = Choice ~ .

tuned.linear = tune.svm(svm_form, data = book.train.split,
                 kernel = "linear",
                 gamma = seq(.01, .1, by = .01), 
                 cost = seq(.1, 1, by = .1))
```

```{r}
mysvm.linear = svm(formula = svm_form, 
            data = book.train.split, 
            gamma =tuned.linear$best.parameters$gamma, 
            cost = tuned.linear$best.parameters$cost)

summary(mysvm.linear)

```

```{r}
# Predict on the test split
predict.test.linear = predict(mysvm.linear, 
                     book.test.split, 
                     type = "response")

caret::confusionMatrix(predict.test.linear, book.test.split$Choice, positive = "1")

```


```{r}
predict.test.linear = predict(mysvm.linear, 
                     BBBC_Test_SVM, 
                     type = "response")

caret::confusionMatrix(predict.test.linear, BBBC_Test_SVM$Choice, positive = "1")

```




#### SVM with Radial Kernel

```{r}
svm_form = Choice ~ .

tuned.radial = tune.svm(svm_form, data = book.train.split, 
                 gamma = seq(.01, .1, by = .01), 
                 cost = seq(.1, 1, by = .1))
```


```{r}
mysvm.radial = svm(formula = svm_form, 
            data = book.train.split, 
            gamma =tuned.radial$best.parameters$gamma, 
            cost = tuned.radial$best.parameters$cost)

summary(mysvm.radial)

```

```{r}
# Predict on the test split
svmpredict.radial = predict(mysvm.radial, 
                     book.test.split, 
                     type = "response")

caret::confusionMatrix(svmpredict.radial, book.test.split$Choice, positive = "1")

```

```{r}
predict.test.radial = predict(mysvm.radial, 
                     BBBC_Test_SVM, 
                     type = "response")

caret::confusionMatrix(predict.test.radial, BBBC_Test_SVM$Choice, positive = "1")

```





#### SVM with Polynomial Kernel

```{r}
svm_form = Choice ~ .

tuned.poly = tune.svm(svm_form, data = book.train.split,
                      kernel = "polynomial",
                      degree = seq(2, 4, by = 1), 
                      coef0 = c(0.01, 0.1, 0.25, 0.5, 0.75, 1, 2, 3), 
                      cost = seq(.1, 1, by = .1))
```


```{r}
mysvm.poly = tuned.poly$best.model

summary(mysvm.poly)

```

```{r}
# Predict on the test split
svmpredict.poly = predict(mysvm.poly, 
                     book.test.split, 
                     type = "response")

caret::confusionMatrix(svmpredict.poly, book.test.split$Choice, positive = "1")

```

```{r}
predict.test.poly = predict(mysvm.poly, 
                     BBBC_Test_SVM, 
                     type = "response")

caret::confusionMatrix(predict.test.poly, BBBC_Test_SVM$Choice, positive = "1")

```


